Automatic Metadata Generation for Fish Specimen Image Collections {\footnotesize \thanks{Research supported by NSF OAC \#1940233 and \#1940322.} }

### 2021-04-05 paper/paper_new.tex
Updates from Overleaf

Digitized specimen accessible in open access repositories provide a rich, extensive data source machine-learning and scientific discovery, however, these resources remain largely untapped due image quality and metadata limitations. The overall goal is addressing this need by combining machine-learning and image informatics techniques. The overall goal of our

%Following is original Goals/objectives text: The amount of data is immense, with image collections containing 10,000's of entries. These discoveries will only be possible with computational techniques that can process this vast amount of data.

@@

%The automated metadata generation methods for our project were developed to work on a specific set of images, the INHS\ Fish Collection~\cite{INHS}.  Most of these images have been configured, produced and acquired with a standard procedure.  The images used for our study contain one fish placed on a bright, white background and contain an information tag and the same ruler.  See Figure \ref{fig:teaser} for two example images from the collection. While training and focusing our system on images with very similar compositions and visual properties may limit its immediate applicability, our efforts do demonstrate the potential that machine learning and image informatics techniques have for automatically generating metadata for biological specimen image collections.

@@

Table \ref{tab:dataset} lists the number of instances for each class 
used in our training dataset.
All of the training data was labeled by hand using \verb|makesense.ai| \cite{make-sense}
% While the ultimate goal of the project is to generalize to images from a variety of specimen collections, we have thus far focused only
on images from the INHS\ Fish Collection \cite{INHS}.
Using \verb|detectron|'s default training scheme, the model was trained for \(100,000\) epochs. All instance types were included in a single object detection model. The total cross--entropy loss began at \(2.948\) and ended at \(0.158\) for the training data.

@@

Final: {"data_time": 0.006645713496254757, "eta_seconds": 0.0, "fast_rcnn/cls_accuracy": 0.984375,
"fast_rcnn/false_negative": 0.0, "fast_rcnn/fg_cls_accuracy": 1.0, "iteration": 99999,
"loss_box_reg": 0.054501257836818695, "loss_cls": 0.03908616118133068, "loss_mask": 0.06379510834813118,
"loss_rpn_cls": 0.0010549294238444418, "loss_rpn_loc": 0.005114196799695492, "lr": 0.02,
"mask_rcnn/accuracy": 0.9722937165937804, "mask_rcnn/false_negative": 0.012800597698663846,
"mask_rcnn/false_positive": 0.05609636495338312, "roi_head/num_bg_samples": 115.75,
"roi_head/num_fg_samples": 12.25, "rpn/num_neg_anchors": 246.5, "rpn/num_pos_anchors": 9.5,
"time": 0.3563086399990425, "total_loss": 0.15775199318886735}
\end{verbatim}
\end{comment}
\subsection{Pixel Analysis}
The masks and bounding boxes produced by \verb|detectron| are generally quite good, although they almost never completely or tightly enclose the
detected objects.
This is problematic for the detected fish objects in our analyzed images,
where the most accurate segmentation is desired.
The mask may include additional background as part of the fish, or the
bounding box may clip away part(s) of the fish. To solve these
shortcomings, we utilize pixel analysis methods commonly found in image
informatics to produce more accurate objects masks and bounding boxes.

@@

\subsubsection{Adjusting the Bounding Box}
With an accurate mask generated, it is then necessary to check whether the bounding box needs expanded or shrunk along any edges. Expansion is done first, buy checking whether any edge intersects with any of the foreground mask pixels. If one does, it is expanded out by 1 pixel. If any edges are expanded, the whole the process, masking and expansion process is repeated to account for any changes in average intensities. Once no edges contain foreground pixels, the mask is the shrunk. This process is more efficient as all that is required is to shrink each edge by one up until the point they contain one or more foreground pixels. Once the shrinkage step is accomplished, the final mask and bounding box have been generated.

\subsubsection{Fallback} The pixel analysis process occasionally fails,
e.g.~ flood-filling does not produce a large enough blob or
the bounding box adjustment does not terminate.
This can occur if certain flood fill operations behave unexpectedly, or if the image is too washed out or otherwise atypical for the thresholding process to work correctly. In the event this happens, the original mask and bounding box generated by \verb|detectron| is used for metadata generation.

>>>>

Digitized specimen accessible in open access repositories provide a rich, extensive data source machine-learning and scientific discovery, however, these resources remain largely untapped due image quality and metadata limitations. The overall goal is addressing this need by combining machine-learning and image informatics techniques. The overall goal of our research is to offer a machine-driven [*novel/innovative?? blah...] approach to current manual metadata generation that is prohibitively costly in both terms of labor and time, and to further offer a [*novel] approach to computing higher-level metadata that supports scientific inquiry. Three key objectives are to:
1) Test [*Explore, Demonstrate the us of] Facebook's AI Research's dectectron tool
Specific aims are to use dectectron to identify five object classes: fish, fish eyes, rulers, and the numbers 2 and 3 on rulers, as shown in Figure 2.
2) Pursue [*Investigate] image processing at the pixel level.
Pilot testing found the dectectron bounding boxes could be more tightly enclosed the detected objects. We will see if additional masking background as part of the fish, or the
bounding box may clip away part(s) of the fish with pixel analysis methods commonly
found in image informatics. Specific aim [*objectives] of our research are to:
%not sure we have to list all the details, but something like the below, with the 4 aspects of pixel analysis we tested/investigate/experimented with 
1. Identify the appropriate threshold. We will compute optimal scaling between [*or is it Otsu thresholding] to identify an improved threshold value. [Or we will determine is we can computed as the halfway point between the Otsu threshold value and the mean of the background intensities.]
2. Consolidate the foreground
3. Adjusting the bounding box. [We will use computational methods for masking [calcuate a mask?] to determine if bounding box needs to be shrunk or expanded]
% Jane confused here, if this belongs here, or below
4. Developing a fallback approach for when flood fill operations behave unexpectedly the image is too washed out.
3) Compute [Test our ability to calculate] the following higher-level metadata properties: 
1) contrast (intensities of
the foreground and background, 2) centroid and eye center, and if eye is found, the fish 3) side 4) primary axis and clock value, finally the 5) scale and length.
%Following is original Goals/objectives text: The amount of data is immense, with image collections containing 10,000's of entries. These discoveries will only be possible with computational techniques that can process this vast amount of data.

@@

The automated metadata generation methods for our project were developed to work on a specific set of images, the INHS\ Fish Collection~\cite{INHS}.  Most of these images have been configured, produced and acquired with a standard procedure.  The images used for our study contain one fish placed on a bright, white background and contain an information tag and the same ruler.  See Figure \ref{fig:teaser} for two example images from the collection. While training and focusing our system on images with very similar compositions and visual properties may limit its immediate applicability, our efforts do demonstrate the potential that machine learning and image informatics techniques have for automatically generating metadata for biological specimen image collections.

@@

Table \ref{tab:dataset} lists the number of instances for each class 
used in our training dataset.
All of the training data was labeled by hand using \verb|makesense.ai| \cite{make-sense}
% While the ultimate goal of the project is to generalize to images from a variety of specimen collections, we have thus far focused only
on images from the INHS\ Fish Collection \cite{INHS}.
Using \verb|detectron|'s default training scheme, the model was trained for \(100,000\) epochs. All instance types were included in a single object detection model. The total cross-entropy loss began at \(2.948\) and ended at \(0.158\) for the training data.

@@

Final: {"data_time": 0.006645713496254757, "eta_seconds": 0.0, "fast_rcnn/cls_accuracy": 0.984375,
"fast_rcnn/false_negative": 0.0, "fast_rcnn/fg_cls_accuracy": 1.0, "iteration": 99999,
"loss_box_reg": 0.054501257836818695, "loss_cls": 0.03908616118133068, "loss_mask": 0.06379510834813118,
"loss_rpn_cls": 0.0010549294238444418, "loss_rpn_loc": 0.005114196799695492, "lr": 0.02,
"mask_rcnn/accuracy": 0.9722937165937804, "mask_rcnn/false_negative": 0.012800597698663846,
"mask_rcnn/false_positive": 0.05609636495338312, "roi_head/num_bg_samples": 115.75,
"roi_head/num_fg_samples": 12.25, "rpn/num_neg_anchors": 246.5, "rpn/num_pos_anchors": 9.5,
"time": 0.3563086399990425, "total_loss": 0.15775199318886735}
\end{verbatim}
\end{comment}
\subsection{Pixel Analysis}
The masks and bounding boxes produced by \verb|detectron| are generally quite good, although they almost never completely or tightly enclose the
detected objects.
This is problematic for the detected fish objects in our analyzed images,
where the most accurate segmentation is desired.
The mask may include additional background as part of the fish, or the
bounding box may clip away part(s) of the fish. To solve these
shortcomings, we utilize pixel analysis methods commonly found in image
informatics to produce more accurate object masks and bounding boxes.

@@

\subsubsection{Adjusting the Bounding Box}
With an accurate mask generated, it is then necessary to check whether the bounding box needs expanded or shrunk along any edges. Expansion is done first, by checking whether any edge intersects with any of the foreground mask pixels. If one does, it is expanded out by 1 pixel. If any edges are expanded, the whole process of masking and expansion process is repeated to account for any changes in average intensities. Once no edges contain foreground pixels, the mask is then shrunk. This process is more efficient as all that is required is to shrink each edge by one up until the point they contain one or more foreground pixels. Once the shrinkage step is accomplished, the final mask and bounding box have been generated.

\subsubsection{Fallback} The pixel analysis process occasionally fails,
e.g.\ when flood-filling does not produce a large enough blob or
the bounding box adjustment does not terminate.
This can occur if certain flood fill operations behave unexpectedly, or if the image is too washed out or otherwise atypical for the thresholding process to work correctly. In the event this happens, the original mask and bounding box generated by \verb|detectron| is used for metadata generation.
